Analyzing Security-Cost Trade-off for a \\ Fully Homomorphic Encryption Scheme

### 2013-06-16 Report/article.tex

\documentclass{acm_proc_article-sp}
\makeatletter
\newif\if@restonecol
\makeatother
\let\algorithm\relax
\let\endalgorithm\relax
\usepackage{hyperref}
\usepackage[all]{hypcap}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{listings}
\usepackage{algorithm2e}
%\usepackage{amsfont}
%\usepackage{amsmath}
%\usepackage{amssymb}
\hypersetup{
  bookmarks=true,         % show bookmarks bar?
  unicode=false,          % non-Latin characters in Acrobat’s bookmarks
  pdftoolbar=true,        % show Acrobat’s toolbar?
  pdfmenubar=true,        % show Acrobat’s menu?
  pdffitwindow=false,     % window fit to page when opened
  pdfstartview={FitH},    % fits the width of the page to the window
  pdftitle={My title},    % title
  pdfauthor={Author},     % author
  pdfsubject={Subject},   % subject of the document
  pdfcreator={Creator},   % creator of the document
  pdfproducer={Producer}, % producer of the document
  pdfkeywords={keyword1} {key2} {key3}, % list of keywords
  pdfnewwindow=true,      % links in new window
  colorlinks=true,       % false: boxed links; true: colored links
  linkcolor=cyan,          % color of internal links (change box color with linkbordercolor)
  citecolor=green,        % color of links to bibliography
  filecolor=magenta,      % color of file links
  urlcolor=cyan           % color of external links
}

\usepackage{listings}
\usepackage{algorithm2e}
\usepackage{verbatim}
\begin{document}
\title{Analyzing Security-Cost Trade-off for a \\ Fully Homomorphic Encryption Scheme}

@@

1. \textbf{XOR of an $n$-bit sequence}\\
2. \textbf{Sum of :}  two $n$ bit sequences  \newline 
3. \textbf{Majority bit of an $n$-bit sequence} \\
4. \textbf{Product of two $n\times n$ boolean matrices} \\
5. \textbf{Sorting of $n$ sequences of $nbits$ using :} \newline\noindent
\phantom{x}\hspace{3ex} a) Insertion sort \phantom{x}\hspace{2ex} b) Bitonic sort \\
	\phantom{x}\hspace{3ex} 	     c) Odd-Even Merge sort

@@

\textbf{XOR of $n$-bits :} \texttt{XOR} of $n$-bits requires $n-1$ \texttt{XOR} gates. \newline \newline
\textbf{Sum of two $n$-bit sequences :} Sum of two $n$-bits sequences is performed using $n$ \texttt{full-adder} circuits (provided as \texttt{fhe\_fulladd(...)} function in the Scarab library). The cost of a full adder circuit is $2$ \texttt{XOR} to calculate the sum and $2$ \texttt{AND}, $1$ \texttt{XOR} and $1$ \texttt{OR} for the carry. Hence the total complexity is $3n$ \texttt{AND}, $5n$ \texttt{XOR}. \newline \newline
\textbf{Majority of $n$-bits :} Majority bit is evaluated by obtaining the sum of the bits and then returning the result of the comparison of the sum and $n/2$. This comparison is done using algorithm \autoref{Code:algo}. The operation is interesting in the sense that if the same program is evaluated on the user's side, he has to store the sum of the bits, hence  needs $\mathrm{\lfloor\log_{2}n \rfloor +1}$ bits of memory. However, if he performs the operation remotely on a cloud, he only requires a constant number of bits to retrieve the result. \newline \newline
\textbf{Product of two $n\times n$ matrices :} This requires $n^3$ \texttt{AND} gates and \texttt{$n^2(n-1)$} \texttt{XOR} gates. \newline \newline
\textbf{Sorting :} Several straight-line programs for sorting have have been proposed in literature \cite{dk}, \cite{Batcher:1968:SNA:1468075.1468121}. These programs use a comparator gate : a logical gate with two inputs and two outputs, that computes the minimum and maximum of two $n$ bit sequences. Algorithm \autoref{Code:algo} implements such a comparator gate. We do not need to perform any transformation for these algorithms. The algorithms used for Bitonic and Odd-Even merge sort are provided in the appendix. They accept $n$ which are powers of $2$.
\newline \newline
\textbf{Insertion sort :} When allowing for parallel comparators, bubble sort and insertion sort are identical. Hence $n(n-1)/2$ comparator gates are used for insertion sort.\newline \newline
\textbf{Bitonic Sort :} \autoref{bitsort} and \autoref{bitmerge} are used to perform Bitonic Sort. The algorithm uses $\mathrm{log}(n)·(\mathrm{log}(n)+1)/2$ comparator gates. \newline \newline
\textbf{Odd-Even Merge sort :} \autoref{oesort} and \autoref{oemerge} are used to perform Odd-Even merge sort and hence  $n(\mathrm{log}^{2}(n)-1)/2 + 1 $ comparator gates are required.\newline
We note that the sorting programs used do not increase the size of the circuit operating on encrypted input bit sequences.

@@

\begin{figure}[t] %on ouvre l'environnement figure
\centering
\includegraphics[width=7cm]{fsort1.pdf} 
\caption{Run time of KeyGen(ms)} 
\label{fig:image_f2} %l'étiquette pour faire référence à cette image

\centering
\includegraphics[width=7cm]{fsort2.pdf} 
\caption{Run time of KeyGen(ms)} 
\label{fig:image_f2} %l'étiquette pour faire référence à cette image
\centering
\includegraphics[width=7cm]{fsort3.pdf} 
\caption{Run time of KeyGen(ms)} 
\label{fig:image_f2} %l'étiquette pour faire référence à cette image
\end{figure}

\begin{figure}[t] %on ouvre l'environnement figure
\centering
\includegraphics[width=7cm]{fsort4.pdf} 
\caption{Run time of KeyGen(ms)} 
\label{fig:image_f2} %l'étiquette pour faire référence à cette image
\end{figure}

\begin{figure}[t] %on ouvre l'environnement figure
\centering
\includegraphics[width=7cm]{fsort5.pdf} 
\caption{Run time of KeyGen(ms)} 
\label{fig:image_f2} %l'étiquette pour faire référence à cette image
\end{figure}

\begin{figure}[t] %on ouvre l'environnement figure
\centering
\includegraphics[width=7cm]{fsort6.pdf} 
\caption{Run time of KeyGen(ms)} 
\label{fig:image_f2} %l'étiquette pour faire référence à cette image
\end{figure}

\begin{figure}[t] %on ouvre l'environnement figure
\centering
\includegraphics[width=7cm]{fsort7.pdf} 
\caption{Run time of KeyGen(ms)} 
\label{fig:image_f2} %l'étiquette pour faire référence à cette image
\end{figure}

\begin{figure}[t] %on ouvre l'environnement figure
\centering
\includegraphics[width=7cm]{fsort8.pdf} 
\caption{Run time of KeyGen(ms)} 
\label{fig:image_f2} %l'étiquette pour faire référence à cette image
\end{figure}

@@

Intially, the cryptographic parameters were set to $(N = 8, \nu = 2^{384}, \mu = 4, s_1 = 8, s_2 = 5)$ which is the default setting in the software. The \textcolor{orange}{above figures} show the variation of CPU time and Wall time with the increase in the number of sequences to sort and the number of bits. These experiments were done for Insertion Sort algorithm. We also present a  curve fitting plot that verifies the quadratic complexity of this algorithm.

@@

\begin{tabular}{|l|l|l||}
  \hline
  (Number of sequences,  & Maximum time (in ms) & Minimum time (in ms) \\
  Number of bits) &  &    \\
  \hline
  (4, 4) &  27958 & 27455\\
  (8, 4)  &    108197  & 103683\\
  (16,4) &   354188 &  341522\\
  (32,4) &  1587427 &  1044640\\
  (32,6) &   1629722 & 610464\\
\hline
\end{tabular}
\label{OddEvensort2}
}
\end{table}

When we increase the security of the system by setting the parameters to  $(N = 256, \nu = 16, \mu = 2, s_1 = 5000, s_2 = 5)$(as suggested in \cite{cryptoeprint:2009:571}) which ensures $25$ bits of security, the Odd-Even sorting algorithm on  $4$ sequences, each of $4$ bits takes hours. By lowering the security level to only $8$ bits using the parameters :$(N = 16, \nu = 32, \mu = 2, s_1 = 85, s_2 = 5)$, the result did not decrypt to the desired output; which we believe is an implementation bug in the bootstrapping mechanism. However, by increasing $\mu$ also increases security, and setting the parameters to :  $(N = 8, \nu = 2^{384}, \mu = 8, s_1 = 8, s_2 = 5)$ we obtained the results as in \autoref{OddEvensort2} :
\section{Conclusion}

>>>>

\documentclass{acm_proc_article-sp}
\makeatletter
\newif\if@restonecol
\makeatother
\let\algorithm\relax
\let\endalgorithm\relax
\usepackage{hyperref}
\usepackage[all]{hypcap}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{listings}
\usepackage{algorithm2e}
\usepackage{subfigure}
%\usepackage{amsmath}
%\usepackage{amssymb}
\hypersetup{
  bookmarks=true,         % show bookmarks bar?
  unicode=false,          % non-Latin characters in Acrobat’s bookmarks
  pdftoolbar=true,        % show Acrobat’s toolbar?
  pdfmenubar=true,        % show Acrobat’s menu?
  pdffitwindow=false,     % window fit to page when opened
  pdfstartview={FitH},    % fits the width of the page to the window
  pdftitle={My title},    % title
  pdfauthor={Author},     % author
  pdfsubject={Subject},   % subject of the document
  pdfcreator={Creator},   % creator of the document
  pdfproducer={Producer}, % producer of the document
  pdfkeywords={keyword1} {key2} {key3}, % list of keywords
  pdfnewwindow=true,      % links in new window
  colorlinks=true,       % false: boxed links; true: colored links
  linkcolor=cyan,          % color of internal links (change box color with linkbordercolor)
  citecolor=green,        % color of links to bibliography
  filecolor=magenta,      % color of file links
  urlcolor=cyan           % color of external links
}
\usepackage{listings}
\usepackage{algorithm2e}
\usepackage{verbatim}
\begin{document}
\title{Analyzing Security-Cost Trade-off for a \\ Fully Homomorphic Encryption Scheme}

@@

1. \textbf{XOR of an $n$-bit sequence}\\
2. \textbf{Sum of :}  two $n$ bit sequences  \newline 
3. \textbf{Majority bit of an $n$-bit sequence} \\
4. \textbf{Product of two $n\times n$ boolean matrices} \\
5. \textbf{Sorting of $n$ sequences of $nbit$ using :} \newline\noindent
\phantom{x}\hspace{3ex} a) Insertion sort \phantom{x}\hspace{2ex} b) Bitonic sort \\
	\phantom{x}\hspace{3ex} 	     c) Odd-Even Merge sort

@@

\textbf{XOR of $n$-bits :} \texttt{XOR} of $n$-bits requires $n-1$ \texttt{XOR} gates. \newline \newline
\textbf{Sum of two $n$-bit sequences :} Sum of two $n$-bits sequences is performed using $n$ \texttt{full-adder} circuits (provided as \texttt{fhe\_fulladd(...)} function in the Scarab library). The cost of a full adder circuit is $2$ \texttt{XOR} to calculate the sum and $2$ \texttt{AND}, $1$ \texttt{XOR} and $1$ \texttt{OR} for the carry. Hence the total complexity is $3n$ \texttt{AND}, $5n$ \texttt{XOR}. \newline \newline
\textbf{Majority of $n$-bits :} Majority bit is evaluated by obtaining the sum of the bits and then returning the result of the comparison of the sum and $n/2$. This comparison is done using algorithm \autoref{Code:algo}. The operation is interesting in the sense that if the same program is evaluated on the user's side, he has to store the sum of the bits, hence  needs $\mathrm{\lfloor\log_{2}n \rfloor +1}$ bits of memory. However, if he performs the operation remotely on a cloud, he only requires a constant number of bits to retrieve the result. \newline \newline
\textbf{Product of two $n\times n$ matrices :} This requires $n^3$ \texttt{AND} gates and \texttt{$n^2(n-1)$} \texttt{XOR} gates. \newline \newline
\textbf{Sorting :} Several straight-line programs for sorting have have been proposed in literature \cite{dk}, \cite{Batcher:1968:SNA:1468075.1468121}. These programs use a comparator gate : a logical gate with two inputs and two outputs, that computes the minimum and maximum of two $n$ bit sequences. Algorithm \autoref{Code:algo} implements such a comparator gate. We do not need to perform any transformation for these algorithms. The algorithms used for Bitonic and Odd-Even merge sort are provided in the appendix. They accept $n$ which are powers of $2$.
\newline \newline
\textbf{Insertion sort :} When allowing for parallel comparators, bubble sort and insertion sort are identical. Hence $n(n-1)/2$ comparator gates are used for insertion sort.\newline \newline
\textbf{Bitonic Sort :} \autoref{bitsort} and \autoref{bitmerge} are used to perform Bitonic Sort. The algorithm uses $n\mathrm{log}(n)·(\mathrm{log}(n)+1)/4$ comparator gates. \newline \newline
\textbf{Odd-Even Merge sort :} \autoref{oesort} and \autoref{oemerge} are used to perform Odd-Even merge sort and hence  $n(\mathrm{log}^{2}(n)-1)/2 + 1 $ comparator gates are required.\newline
We note that the sorting programs used do not increase the size of the circuit operating on encrypted input bit sequences.

@@

\begin{figure}[!H]
\centering
\includegraphics[width=7cm]{fsort3.pdf} 
\caption{Runtime for Insertion sort, nbit=4} 
\label{fig:image_sf3} %l'étiquette pour faire référence à cette image
\end{figure}

\begin{figure}[!H] %on ouvre l'environnement figure
\centering
\includegraphics[width=7cm]{fsort4.pdf} 
\caption{Runtime for Insertion sort, nbit=6} 
\label{fig:image_sf4} %l'étiquette pour faire référence à cette image
\end{figure}

\begin{figure}[!H] %on ouvre l'environnement figure
\centering
\includegraphics[width=7cm]{fsort5.pdf} 
\caption{Run time for Insertion sort, nbit=8} 
\label{fig:image_sf5} %l'étiquette pour faire référence à cette image
\end{figure}

\begin{figure}[!H] %on ouvre l'environnement figure
\centering
\includegraphics[width=7cm]{fsort6.pdf} 
\caption{CPU time for Insertion sort, nbit=4} 
\label{fig:image_sf6} %l'étiquette pour faire référence à cette image
\end{figure}

\begin{figure}[!H] %on ouvre l'environnement figure
\centering
\includegraphics[width=7cm]{fsort8.pdf} 
\caption{CPU time for Insertion sort, nbit=6} 
\label{fig:image_sf8} %l'étiquette pour faire référence à cette image
\end{figure}

\begin{figure}[!H]%on ouvre l'environnement figure
\centering
\includegraphics[width=7cm]{fsort7.pdf} 
\caption{CPU time for Insertion sort, nbit=8} 
\label{fig:image_sf7} %l'étiquette pour faire référence à cette image
\end{figure}

\begin{figure}[!H]
\centering
\includegraphics[width=7cm]{fsort2.pdf} 
\caption{Runtime Comparison for insertion sort} 
\label{fig:image_sf2} %l'étiquette pour faire référence à cette image
\end{figure}

\begin{figure}[!H]%nouvre l'environnement figure
\centering
\includegraphics[width=7cm]{fsort1.pdf} 
\caption{Curve fitting for insertion sort}
\label{fig:image_sf1} %l'étiquette pour faire référence à cette image
\end{figure}

@@

Intially, the cryptographic parameters were set to $(N = 8, \nu = 2^{384}, \mu = 4, s_1 = 8, s_2 = 5)$ which is the default setting in the software. \autoref{fig:image_sf3},\autoref{fig:image_sf4}, \autoref{fig:image_sf5}, \autoref{fig:image_sf6}, \\ \autoref{fig:image_sf8}, \autoref{fig:image_sf7} and \autoref{fig:image_sf2} show the variation of CPU time and Wall time with the increase in the number of sequences to sort and the number of bits. These experiments were done for Insertion Sort algorithm. \autoref{fig:image_sf1} presents a  curve fitting plot that verifies the quadratic complexity of this algorithm.

@@

\begin{table}
\caption{Wall time for Odd-Even Merge Sort with increased security}
\resizebox{8cm}{!}{

\begin{tabular}{|l|l|l||}
  \hline
  (Number of sequences,  & Maximum time (in ms) & Minimum time (in ms) \\
  Number of bits) &  &    \\
  \hline
  (4, 4) &  27958 & 27455\\
  (8, 4)  &    108197  & 103683\\
  (16,4) &   354188 &  341522\\
  (32,4) &  1587427 &  1044640\\
  (32,6) &   1629722 & 1610464\\
\hline
\end{tabular}
\label{OddEvensort2}
}
\end{table}

When we increase the security of the system by setting the parameters to  $(N = 256, \nu = 16, \mu = 2, s_1 = 5000, s_2 = 5)$(as suggested in \cite{cryptoeprint:2009:571}) which ensures $25$ bits of security, the Odd-Even sorting algorithm on  $4$ sequences, each of $4$ bits takes hours. By lowering the security level to only $8$ bits using the parameters :$(N = 16, \nu = 32, \mu = 2, s_1 = 85, s_2 = 5)$, the result did not decrypt to the desired output; which we believe is an implementation bug in the bootstrapping mechanism. Lowering the security even further to only $5$ bits using the parameters: $(N = 8, \nu = 2^8, \mu = 4, s_1 = 65, s_2 = 5)$, and testing odd-even sorting for $32$ sequences of $6$ bits each took more than an hour. However, increasing $\mu$ also increases security, and setting the parameters to :  $(N = 8, \nu = 2^{384}, \mu = 8, s_1 = 8, s_2 = 5)$ we obtained the results as in \autoref{OddEvensort2}. The results show that this set of parameters provide a better security and at almost no extra cost. For this result to be conclusive, we need to perfom tests for large $n$.
\section{Conclusion}
